Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-05-17 main.tex
cleanup

\noindent
The symbolic-weighted transducer~$T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
\[
\begin{array}{rccl}
\weight_T(q, \epsilon, \epsilon, q') & = &              \wei(q, \epsilon, \epsilon, q')\\
\weight_T(q, s, t, q') & = & 
             \displaystyle\bigoplus_{q'' \in Q} &
             \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
 & \oplus &  \displaystyle\bigoplus_{q'' \in Q} &
             \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
 & \oplus &  \displaystyle\bigoplus_{\begin{array}{c}
                                     \scriptstyle q'' \in Q\\[-2pt]
                                     \scriptstyle s = au, a \in \Sigma
                                     \end{array}} &
   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
\end{array}
\]
%
We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
%
The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
%
%** The application of $\oplus$ can be seen as a ND choice between transitions **

@@

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \epsilon, q_1) & = & \alpha\\
\end{array}
\]
%
Finally, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switch from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
\endex
\end{example}

@@

\noindent
The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
can be synthesized as above into a function
$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

@@

>>>>

\noindent
The symbolic-weighted transducer~$T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
\[
\begin{array}{rccl}
\weight_T(q, \epsilon, \epsilon, q') & = &              \wei(q, \epsilon, \epsilon, q')\\
\weight_T(q, s, t, q') & = & 
             \displaystyle\bigoplus_{q'' \in Q} &
             \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
 & \oplus &  \displaystyle\bigoplus_{q'' \in Q} &
             \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
 & \oplus &  \displaystyle\bigoplus_{\begin{array}{c}
                                     \scriptstyle q'' \in Q\\[-2pt]
                                     \scriptstyle s = au, a \in \Sigma
                                     \end{array}} &
   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
\end{array}
\]
%
We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
%
The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
%
%** The application of $\oplus$ can be seen as a ND choice between transitions **
% choose to advance head in input, in output or both (or none)

@@

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \epsilon, q_1) & = & \alpha\\
\end{array}
\]
%
Finally, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switch from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
\endex
\end{example}

@@

\noindent
As above in the case of \SWT, 
the transition functions $\wei_\epsilon$ and $\wei_\Sigma$
can be synthesized into a single function
$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

@@

%\input{transcription}

### 2021-05-17 transcription.tex
cleanup

\subsection{Application to Automated Music Transcription}
\label{sec:transcription}
Symbolic Automated Music Transcription
and analysis of music performances

@@

A music performance is a finite sequence of events in a set~$\Sigma$.
Every event $e \in \Sigma$ has attributes such from a finite domain, 
like a number of key for a note 
or a flag indicating that it is a rest 
(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
and a velocity value (0..127 in~\cite{SMF})).
%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
Moreover, it contains a RTU value $\ioi{e}$ (real number) 
representing the time distance to the next event, 
or to the end of performance for the last event,
also called \emph{inter-onset interval}.

@@

We consider an alphabet $\Delta$, every symbol of which is 
composed of a tag, in a finite set $\Xi$, 
and an MTU (rational) IOI duration value.
%The alphabet $\Delta$ 
It is partitioned into 
$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
like in Section~\ref{section:SWVPA}.
%
\noindent
The symbols of $\Deltai$ represent events:
% (infinite alphabet of internal symbols) made of:
with tags indicating a new note or grace-note (with null IOI), 
a rest or the continuation of the previous note (tie or dot).
%
The elements of $\Deltac \uplus \Deltar$ are matched
markups for describing the structure of the score, 
\ie the hierarchical grouping of events, and also, 
importantly the division of time in measures, tuplets...
%- parentheses for time divisions : tuplets, bars...
(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
They contain additional info such as tuple number, beaming policy...

@@

>>>>

% !TEX root = main.tex
%
% Application to AMT
%
\subsection{Application to Automated Music Transcription}
\label{sec:transcription}
Symbolic Automated Music Transcription
and analysis of music performances

@@

A music performance is a finite sequence of events in a set~$\Sigma$.
Every event $e \in \Sigma$ has attributes such from a finite domain, 
like a number of key for a note 
or a flag indicating that it is a rest 
(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
and a velocity value (0..127 in~\cite{SMF})).
%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
Moreover, it contains a time value $\ioi{e}$ (real number) in RTU (seconds)
representing the time distance to the next event, 
or to the end of performance for the last event,
also called \emph{inter-onset interval}.

@@

We consider an alphabet $\Delta$, every symbol of which is 
composed of a tag, in a finite set $\Xi$, 
and an IOI duration value expressed in MTU (rational number).
%The alphabet $\Delta$ 
It is partitioned into 
$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
like in Section~\ref{section:SWVPA}.
%
\noindent
The symbols of $\Deltai$ represent events:
% (infinite alphabet of internal symbols) made of:
with tags indicating a new note or grace-note (with null IOI), 
a rest or the continuation of the previous note (tie or dot).
%
The elements of $\Deltac \uplus \Deltar$ are matched
markups for describing the structure of the score, 
\ie the hierarchical grouping of events, and also, 
importantly the division of time in measures, tuplets...
%- parentheses for time divisions : tuplets, bars...
(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
They contain additional info such as tuple number, beaming policy...

@@

\subsection{Implementation}

\textsf{qparse} based on a similar approach.

slightly different models of symbolic weighted tree automata
...WTA  specification of a prior score language.

% we hope to gain flexibility with the present framework and extend application field
% from transicription to other problems related to music notation and symbolic recorded performances.
